Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-02 xrb2/paper.tex
add more citations

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\enucdot}{\dot{e}_{nuc}}

@@

We investigate the effect of rotation rate on the flame. With the exception of (citation to the weird 11 Hz flame), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}. There are a number of factors that could explain this lack of observations below 200 Hz. It could be that there is some physical process which inhibits the flame ignition and/or spread at lower rotation rates. It could be that burst oscillations at lower rotation rates are smaller in amplitude and therefore more difficult to detect. It could be that it's not got anything to do with the flame at all, but that neutron stars in accreting low mass X-ray binaries rarely have rotation rates below 200 Hz \MarginPar{look into this?}.

@@

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
and reach higher burning rates, resulting in an $\epsdot$ increase over time.

@@

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost. The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates. \MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?}. This model burns so strongly that it is only run out to $40$ ms. Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run. After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm. There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ at heights up to $\sim 0.8\times 10^4$ cm. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
\MarginPar{NF: should I include a plot for this or is it ok to just say it?}. 
Figure~\ref{fig:4e8_stacked_abar} plots $\bar{A}$ for $\tt{F500\_4E8}$, and similarly shows burning that extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}). A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry. %Can also add a note here about how high abar values are reached across the domain, maybe don or alice can help give a more detailed analysis of abar. Maybe I could also segue into Alice's phase plots

>>>>

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\enucdot}{\dot{e}_{\rm nuc}}

@@

We investigate the effect of rotation rate on the flame. With the exception of (citation to the weird 11 Hz flame \citep{altamirano2010discovery}), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}. There are a number of factors that could explain this lack of observations below 200 Hz. It could be that there is some physical process which inhibits the flame ignition and/or spread at lower rotation rates. It could be that burst oscillations at lower rotation rates are smaller in amplitude and therefore more difficult to detect. It could be that it's not got anything to do with the flame at all, but that neutron stars in accreting low mass X-ray binaries rarely have rotation rates below 200 Hz \MarginPar{look into this?}.

@@

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}\MarginPar{How is the acceleration calculated for this table? I.e. Specifically, at what time / over what time window is it calculated?}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.

@@

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost. The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. \MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?} This model burns so strongly that it is only run out to $40$ ms. Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run. After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm. There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4$ cm. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
\MarginPar{NF: should I include a plot for this or is it ok to just say it?}. 
Figure~\ref{fig:4e8_stacked_abar} plots $\bar{A}$ for $\tt{F500\_4E8}$, and similarly shows burning that extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}). A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry. %Can also add a note here about how high abar values are reached across the domain, maybe don or alice can help give a more detailed analysis of abar. Maybe I could also segue into Alice's phase plots
